---
title: "Structured Abstract"
subtitle: "Time-to-Positivity and Mortality in Bacteremia: Updated Bayesian Meta-Analysis"
date: today
format:
  html:
    fig-width: 10
    toc: false
execute:
  cache: true
  echo: false
  warning: false
  message: false
---

```{r setup}
suppressPackageStartupMessages({
  library(tidyverse)
  library(brms)
  library(metafor)
  library(posterior)
  library(ggdist)
  library(gt)
})
source("R/01_prepare_data.R")
source("R/02_fit_models.R")

es_all  <- prepare_effect_sizes("data/TTP_MetaAnalysis_Extraction_Complete_v4.xlsx",
                                include_inclusive_pool = FALSE,
                                drop_contaminated_ttp  = TRUE)
es_mort <- es_for_outcome(es_all, "mortality")

es_all_inc  <- prepare_effect_sizes("data/TTP_MetaAnalysis_Extraction_Complete_v4.xlsx",
                                    include_inclusive_pool = TRUE,
                                    drop_contaminated_ttp  = TRUE)
es_mort_inc <- es_for_outcome(es_all_inc, "mortality")

fits     <- fit_all_models(es_mort,     cache_dir = "_brms_cache/primary")
fits_inc <- fit_all_models(es_mort_inc, cache_dir = "_brms_cache/sensitivity")
```

```{r summaries}
#| cache: false
source("R/02_fit_models.R")
base_post <- pooled_summary(fits$fit_base)
inc_post  <- pooled_summary(fits_inc$fit_base)
```

## Introduction

Time-to-positivity (TTP) — the interval from blood culture incubation to automated detection of bacterial growth — has been proposed as a prognostic marker in bacteremia. Shorter TTP may reflect higher pathogen burden or virulence, and has been hypothesised to predict mortality. We performed an updated Bayesian random-effects meta-analysis to quantify this association.

## Methods

Five electronic databases were searched through May 2026 (3,196 unique records after deduplication). Studies were eligible if they reported a quantitative TTP–mortality association in adults with bloodstream infection confirmed by automated blood culture monitoring. Log odds ratios were derived from 2×2 tables or back-calculated from reported estimates. A Bayesian hierarchical model with weakly-informative priors was fitted using Stan; heterogeneity was quantified with τ and I². Risk of bias was assessed with the QUIPS tool.

## Results

```{r results-inline}
#| results: asis
n_es      <- nrow(es_mort)
n_studies <- attr(es_all, "n_studies")

cat(sprintf(
  "%d mortality effect sizes from %d studies were pooled (Figure 2).
   Short TTP was associated with higher mortality (OR\u202f=\u202f%.2f, 95%%\u202fCrI\u202f%.2f\u2013%.2f;
   P(OR\u202f>\u202f1)\u202f=\u202f%.0f%%).
   Between-study heterogeneity was substantial (I²\u202f=\u202f%.0f%%, τ\u202f=\u202f%.2f;
   Figure 3).
   The 95%% prediction interval for a new study (%.2f\u2013%.2f) crossed the null,
   reflecting heterogeneity.
   Results were consistent in the %d-study sensitivity pool
   (OR\u202f=\u202f%.2f, 95%%\u202fCrI\u202f%.2f\u2013%.2f).\n",
  n_es, n_studies,
  base_post$median_OR, base_post$ci_OR[1], base_post$ci_OR[2],
  base_post$p_harm * 100,
  base_post$median_i2, base_post$median_tau,
  base_post$pi_OR[1], base_post$pi_OR[2],
  attr(es_all_inc, "n_studies"),
  inc_post$median_OR, inc_post$ci_OR[1], inc_post$ci_OR[2]
))
```

## Conclusions

Short TTP is consistently associated with increased mortality risk in bacteremia across 100 studies; however, substantial between-study heterogeneity limits direct generalisation to individual clinical settings. Prospective studies with pre-specified TTP thresholds are needed to establish clinical utility.

---

### Figure 2 — Forest plot

```{r fig-forest-abstract}
#| label: fig-forest-abstract
#| fig-cap: "Forest plot of OR (short vs long TTP) on mortality — primary pool (n=100 studies). Each row is a single effect size. Pooled Bayesian estimate in red; solid bar = 95% CrI, dashed bar = 95% prediction interval."

fig_h <- max(6, nrow(es_mort) * 0.55 + 2)
knitr::opts_current$set(fig.height = fig_h)

forest_data <- es_mort |>
  arrange(yi) |>
  mutate(
    lbl = paste0(study_id,
                 ifelse(arm_id == "all", "", paste0(" [", arm_id, "]"))),
    lbl = make.unique(lbl, sep = " #"),
    lbl = factor(lbl, levels = lbl),
    or      = exp(yi),
    ci_low  = exp(yi - 1.96 * sei),
    ci_high = exp(yi + 1.96 * sei),
    type = "Study"
  )

pooled_row <- tibble(
  lbl = "POOLED (Bayesian)", or = base_post$median_OR,
  ci_low = base_post$ci_OR[1], ci_high = base_post$ci_OR[2], type = "Pooled"
)
pi_row <- tibble(
  lbl = "POOLED (Bayesian)",
  pi_low  = base_post$pi_OR[1],
  pi_high = base_post$pi_OR[2]
)

forest_data$lbl <- factor(forest_data$lbl,
                           levels = c("POOLED (Bayesian)", levels(forest_data$lbl)))
pooled_row$lbl  <- factor(pooled_row$lbl, levels = levels(forest_data$lbl))
pi_row$lbl      <- factor(pi_row$lbl,     levels = levels(forest_data$lbl))

ggplot(bind_rows(forest_data, pooled_row), aes(x = or, y = lbl, color = type)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point(size = 2.5) +
  geom_errorbar(aes(xmin = pmax(ci_low, 0.05), xmax = pmin(ci_high, 50)),
                width = 0.25, linewidth = 0.7, orientation = "y") +
  geom_errorbar(data = pi_row,
                aes(y = lbl, xmin = pmax(pi_low, 0.05), xmax = pmin(pi_high, 50)),
                width = 0.15, linewidth = 0.7, linetype = "dashed",
                color = "darkred", inherit.aes = FALSE) +
  scale_color_manual(values = c("Study" = "steelblue", "Pooled" = "darkred")) +
  scale_x_log10(breaks = c(0.1, 0.25, 0.5, 1, 2, 4, 8, 16, 32),
                limits  = c(0.05, 50)) +
  labs(subtitle = "Solid red bar = 95% CrI; dashed red bar = 95% prediction interval",
       x = "OR, short vs long TTP (log scale)", y = NULL, color = NULL) +
  theme_minimal(base_size = 9) +
  theme(legend.position  = "bottom",
        axis.text.y      = element_text(size = 7, lineheight = 0.85),
        panel.grid.major.y = element_line(color = "gray92"),
        panel.grid.minor = element_blank())
```

### Figure 3 — Posterior density of pooled OR

```{r fig-posterior-abstract}
#| label: fig-posterior-abstract
#| fig-cap: "Posterior density of the pooled OR (short vs long TTP). The vertical dashed line marks no effect (OR = 1)."
#| fig-height: 4
tibble(OR = base_post$OR_draws) |>
  ggplot(aes(x = OR)) +
  stat_halfeye(fill = "#1f78b4", alpha = 0.6) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  scale_x_continuous(limits = c(0.5, 6), breaks = seq(0.5, 6, 0.5)) +
  labs(x = "Pooled OR (short vs long TTP)", y = "Posterior density") +
  theme_minimal()
```
